Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-31 VMCAI-revised-version/final.tex
up

\begin{proof}
First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $2$ counters.
	This is due to the undecidability of $2$-counter Minsky machine termination.
From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $2$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
such that there exists a terminating run in $M$ for all inputs iff machine $M'$ can reach $q_f'(0,0)$ from any input. 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the two counters until they reach $0$, then, and only then, go in $q_f'$.
Based on this construction, the problem of deciding whether the downward-closed set 
$\{q_f'(0,0)\}$ is reachable in the $2$-counter Minsky machine $M'$
from any input is equivalent to the asking whether there exists a terminating run from every input of machine $M$, that is undecidable~\cite{Min61}. 
Note that by chosing $s_0 \in Q_{M'} \times \N^2 \setminus \{q_f'(0,0) \}$,
deciding whether the downward-closed set 
$ \{q_f'(0,0) \}$ is reachable
from every element of $\post^*(s_0)$ 
is again equivalent to asking whether $M$ terminate on every input.
Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.

% \alain{écrire qqchose ici}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS in which
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
to
a state $ q({\bf 0})$ if and only if
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 the state $q_f(0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed
\end{proof}

>>>>

\begin{proof}
First, note that deciding whether from all inputs, all runs are terminating
 is undecidable for Minsky machines with at least $2$ counters, and, in particular, for deterministic ones as well~\cite{Min61}. Hence 
  deciding whether from all inputs, there exists at least one terminating run
 is undecidable for Minsky machines with at least $2$ counters, since,
 for deterministic Minsky machines, the two problems are one and the same.

Now, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $2$ counters.
%	This is due to the undecidability of $2$-counter Minsky machine structural termination.
From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $2$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
such that there exists a terminating run in $M$ for all inputs iff machine $M'$ can reach $q_f'(0,0)$ from any input. 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the two counters until they reach $0$, then, and only then, go in $q_f'$.
Based on this construction, the problem of deciding whether the downward-closed set 
$\{q_f'(0,0)\}$ is reachable in the $2$-counter Minsky machine $M'$
from any input is equivalent to the asking whether there exists a terminating run from every input of machine $M$, that is undecidable. 
Note that by chosing $s_0 \in Q_{M'} \times \N^2 \setminus \{q_f'(0,0) \}$,
deciding whether the downward-closed set 
$ \{q_f'(0,0) \}$ is reachable
from every element of $\post^*(s_0)$ 
is again equivalent to asking whether $M$ terminate on every input.
Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.

Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS 
which mimick the behaviors of the Minsky machines' counters by using two mirror counters 
for each of them,
and in which
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
	with equal mirror counters
to
a state $ q({\bf 0})$ if and only if
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 the state $q_f(0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
  []
Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed
\end{proof}
